---
title: "day_8_comp"
author: "Julia Parish"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)

```

## Get the lizards data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

```
# Histogram of lizard weights - really bad

```{r}
lizard_weights_histo <- ggplot(data = lizards, 
                               aes(x = weight)) +
                        geom_histogram(fill = "orange", 
                                       color = "purple",
                                       size = 0.5,
                                       linetype = "dotted")
lizard_weights_histo
```

# total_length on x-axis
# weight on y-axis 
# shape (0 - 30)

```{r}
lizard_length <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "yellow",
             color = "red",
             size = 4,
             aplha = 0.5)

lizard_length

```

# Bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length
**this is a bad idea, do not do in future**
```{r}
lizard_bubble <- ggplot(lizards, 
                        aes(x = total_length, y = weight)) +
                geom_point(aes(color = common_name, 
                               size = total_length),
                           shape = 22,
                           fill = "black") +
  theme_update()

lizard_bubble

```

## Facet grid - point plot

```{r}

lizard_wrap <- ggplot(lizards, 
                        aes(x = total_length, y = weight)) +
                geom_point(aes(color = common_name)) +
  theme_update() +
  facet_wrap(~common_name, ncol = 4)

lizard_wrap
```


```{r}
lizard_grid <- ggplot(lizards, 
                        aes(x = total_length, y = weight)) +
                geom_point(aes(color = common_name)) +
  theme_minimal() +
  facet_grid(sex ~ tail)

lizard_grid
```

# Lizard counts by different groups in common_name 

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# OR
lizard_counts <- lizards %>% dplyr::count(common_name)
# also... 2 variables
lizard_counts_cn_tail <- lizards %>% dplyr::count(common_name, tail)

lizard_counts

```

# aes is y first b/c common_name is long and reads better on Y axis
```{r}
liz_count_plot <- ggplot(data = lizard_counts, 
       aes(y = common_name, x = count)) +
  geom_col()

liz_count_plot  
```

# Make common_name a factor

```{r}
liz_count_plot2 <- ggplot(data = lizard_counts, 
       aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Lizard Counts")

liz_count_plot2
```

# Try converting common_name to an ordered factor outside of ggplot

```{r}
lizard_counts2 <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))

class(lizard_counts$common_name)
levels(lizard_counts2$common_name)
```

# Scales - Axis scale breaks, limits & labels
scatterplot: total_length (x) vs weight (y)

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
        geom_point() +
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))


```
# Transform the date column to class Date, then find counts of observations by date

```{r}
lizard_counts_date <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)

class(lizard_counts_date$date)
lizard_counts_date

```

# Make a line plot (geom_line()) of date (x) and count (y)

```{r}
lizards_line <- ggplot(lizard_counts_date, 
                       aes(x = date, y = n)) +
                geom_line() +
                scale_x_date(date_breaks = "3 years",
                             date_labels = "%y")

lizards_line
```



```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  scale_color_gradientn(colors = c("orange", "cyan", "pink", "black")) 

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  scale_color_stepsn(colors = c("green", "blue", "salmon"),
                   breaks = c(0, 20, 40, 50)) 


```

### Update a color schme using a palette in paletter
- make a boxplot with common_name on y axis, total_length on x-axis, with color changing based on common_name
- function to apply recreate common_name (make a new column) tp change class to factor

```{r}
lizard_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

lizard_boxplot <- ggplot(data = lizard_fct, 
                  aes(y = common_name, x = total_length)) +
                  geom_boxplot(aes(fill = common_name),
                               show.legend = FALSE) +
                  scale_fill_paletteer_d(palette = "IslamicArt::jerusalem")
                  
lizard_boxplot
                  
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect("cyan4"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow", size = 15)
    
    #panel.background = element_rect(fill = "yellow",
                                    #color = "purple",
                                    #size = 10),
    #panel.grid.major.x = element_line(color = "red"),
    #panel.grid.major = element_line(color = "red"),
    #panel.grid.minor.y = element_line(color = "orange")
  )


```


