---
title: "day_9_comp"
author: "Julia Parish"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)

```


```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

```

#ggrepel

- Make a subset from lizards, called ww_lizards, that only contains obs for "western whiptail" lizards from the site "sand"

```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", site == "sand")
```

```{r}
ww_lizards_plot <- ww_lizards %>% 
                  ggplot(aes(x = total_length, y = weight)) +
                  geom_point() +
                  geom_text_repel(aes(label = toe_num), size = 3, 
                                  max.overlaps = 20)
                          
ww_lizards_plot 
```

```{r}
# View(gapminder)

# Filter for only Europe, then create a scatterplot of gdpPercap (x) and lifeexp(y) labeled by country name


euro_gdp <- gapminder %>% 
  filter(continent == "Europe", year == 2002) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)

euro_gdp

```

## gghighlight
- want to find toe number 251 - can set a cond't and the data that meets that cond't will be highlights. greys out other obs, labels highlight

```{r}
p <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)

```

# create line plot total_length vs weight, highlight species with weights over 30

```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30)

q
```

## Patchwork for compound figures: have two+ plots into one figure
- can add labels to images/figures
- & themeing gets applied to compound figures
- alternative is calplot

```{r}
p | q
```


```{r}
(p | q) / q &
  theme_minimal()

```

# Explore new graph types

### Marginal plots

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

# unique(whiptail$common_name)

wt <- whiptails %>% 
  ggplot(aes(x = total_length, y = weight)) +
  geom_point() + 
  geom_rug()

wt

```

# Marginal plot w/ box plots

```{r}
liz_1 <- whiptails %>% 
  ggplot(aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) + 
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(liz_1, type = "boxplot", groupColour = TRUE)

liz_1
```
## Beeswarm plot

```{r}
wptails <- whiptails %>% 
  ggplot(aes(x = sex, y = weight)) + 
  geom_beeswarm() + 
  geom_violin(fill = "goldenrod") + 
  geom_boxplot(fill = NA)

wptails

```

## Heatmap with geom_tile
- make a new dataframe called lizard_counts, starting with lizards, with date converted to class Date. Then count by year and common name the number of lizards observed.

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date),
         year = lubridate::year(date))  %>% 
  count(year, common_name)
  
```

```{r}
liz_heatmap <- lizard_counts %>% 
  ggplot(aes(x = year, y = common_name)) +
  geom_tile(aes(fill = n)) + 
  geom_text(aes(label = n), color = "white", size = 3) + 
  scale_fill_viridis_c()

# scale_fill_viridisc() is colorblind friendly!
# scale_fill_gradientn(colors = c("navy", "red", "orange"))

liz_heatmap

```

### Make A MAP!!!! 
- lizard obs are from jornada basin, make a veg map 
- use `read_sf` to read in the "doc.kml" file

```{r}
jornada_vegetation <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()


ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = NA) + 
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Dominant vegetation:") +
  theme_void()


```

